### 2021-06-25 paper.tex
Latex Template for the paper.

\documentclass[sigplan,anonymous,review]{acmart}

\begin{document}
\section{Introduction}
Right now we have the following plan about the structure of the GPCE paper.
\begin{enumerate}
    \item Goal: implementing DSL with dependent type system 
    \item Problem: current two approaches are either doing a deep embedding or
        implementing a typechecker from scratch
    \item Reflection to the rescue: we can use reflection for implementing
        custom compilers 
    \item Our approach by example: demonstrate that this actually works in Agda
        (ps example)
    \item Extracting definitions by pattern matching: matching on the stack
    \item Using normalization to get rid of meta-functions (by using this
        technique, you get macros for free! / or use rewrite rules example:
        exchange o exchange = id)
    \item General technology for writing extractors
    \item This actually scales to realistic and useful examples!
\end{enumerate}

Optional parts:
\begin{enumerate}
    \item Generating assertions
    \item Ensuring termination of recursive functions
\end{enumerate}

\end{document}

### 2021-06-30 paper.tex
Add a bit more structure to the paper.
        NOTE there is no much text in there.  Mostly thoughts
             and parts of the old paper.

@@
 
+%% Bibliography style
+\bibliographystyle{ACM-Reference-Format}
+
+\citestyle{acmauthoryear}   %% For author/year citations
+
+%% Some recommended packages.
+\usepackage{booktabs}   %% For formal tables:
+                        %% http://ctan.org/pkg/booktabs
+\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
+                        %% http://ctan.org/pkg/subcaption
+
+% Add definitions for abbreviations like e.g.; i.e; etc. with
+% correct spacing depending on the parameters.  In this case
+% `all' exposes all the definitions of the package, and `british'
+% makes sure that there is no comma after e.g. or i.e.
+\usepackage[all,british]{foreign}
+
+\usepackage{mathpartir}
+
+\usepackage{todonotes}
+
+\usepackage{varwidth}
+
+\usepackage{agda}
+\usepackage{newunicodechar}
+\usepackage[final]{listings}
+
+\newunicodechar{∷}{::}
+\newunicodechar{→}{\ensuremath{\to}}
+\newunicodechar{ω}{\ensuremath{\omega}}
+\newunicodechar{⊎}{\ensuremath{\uplus}}
+\newunicodechar{≔}{\ensuremath{\coloneqq}}
+\newunicodechar{∎}{\ensuremath{\blacksquare}}
+\newunicodechar{≟}{\ensuremath{\stackrel{?}{=}}}
+\newunicodechar{ᵣ}{\ensuremath{_r}}
+\newunicodechar{ₗ}{\ensuremath{_l}}
+\newunicodechar{▴}{\ensuremath{\blacktriangle}}
+\newunicodechar{▾}{\ensuremath{\blacktriangledown}}
+\newunicodechar{▹}{\ensuremath{\triangleright}}
+\newunicodechar{∸}{\ensuremath{\dot{-}}}
+\newunicodechar{⊛}{\ensuremath{\circledast}}
+\newunicodechar{ᵀ}{\ensuremath{{}^{\text{T}}}}
+\newunicodechar{⌿}{\ensuremath{\not-}}
+
+% Some shortcut commands for agda symbols
+\newcommand{\AD}[1]{\AgdaDatatype{#1}}
+\newcommand{\AC}[1]{\AgdaInductiveConstructor{#1}}
+\newcommand{\AF}[1]{\AgdaFunction{#1}}
+\newcommand{\AB}[1]{\AgdaBound{#1}}
+\newcommand{\AK}[1]{\AgdaKeyword{#1}}
+\newcommand{\AR}[1]{\AgdaField{#1}}
+\newcommand{\AM}[1]{\AgdaModule{#1}}
+\newcommand{\AN}[1]{\AgdaNumber{#1}}
+\newcommand{\AS}[1]{\AgdaString{#1}}
+
+\newcommand\codeblock[1]{%
+  %{\fbox{\begin{varwidth}{0.9\textwidth}#1\end{varwidth}}}
+  {\begin{varwidth}{0.9\textwidth}#1\end{varwidth}}
+}
+
 \begin{document}
+
+%% Title information
+\title[]{Extraction once again}
+
+%% Author with single affiliation.
+\author[A. {\v{S}}inkarovs]{Artjoms {\v{S}}inkarovs}
+\orcid{0000-0003-3292-2985}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                         %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{%
+  \institution{Heriot-Watt University}   %% \institution is required
+  \city{Edinburgh}
+  \state{Scotland}
+  \postcode{EH14 4AS}
+  \country{UK}}
+\email{a.sinkarovs@hw.ac.uk}             %% \email is recommended
+
+%% Author with single affiliation.
+\author[J. Cockx]{Jesper Cockx}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                         %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{%
+  \institution{TU Delft}
+  \city{Delft}
+  \postcode{2628 XE}
+  \country{Netherlands}}
+\email{j.g.h.cockx@tudelft.nl}           %% \email is recommended
+
+%% Abstract
+%% Note: \begin{abstract}...\end{abstract} environment must come
+%% before \maketitle command
+\begin{abstract}
+  Dependently-typed host languages empower users to verify a wide
+  range of properties of embedded languages and
+  programs written in them.  Designers of such embedded languages
+  are faced with a difficult choice between using a shallow or a deep
+  embedding.  The former is easier to use because the entire
+  infrastructure of the host langauge is immediately
+  available. Meanwhile, the latter gives full access to the
+  structure of embedded programs, but is difficult to use in practice,
+  especially when the embedded language is itself dependently typed.
+
+  The main insight presented in this paper is that the choice between
+  shallow and deep embedding can be eliminated by working in a host
+  language with reflection capabilities: we start from a shallow
+  embedding that can use all libraries and tools of the host language,
+  and later use reflection to expose the deep structure of the
+  embedded programs.
+  %
+  Concretely, we apply this technique to embed three
+  programming languages --- Kaleidoscope, SaC, and (a subset of) APL ---
+  into the dependently typed theorem prover Agda, using dependent
+  types to statically enforce several properties of interest.  We then
+  use Agda's reflection capabilities to extract the embedded programs
+  back into the original language, so that the existing toolchain can
+  be leveraged.
+  %
+  In this process, statically verified properties of the host language
+  are mapped onto runtime checks in the target language, allowing
+  extracted programs to interact safely with existing code.
+  %
+  Finally, we demonstrate the feasibility of our approach with the
+  implementation and extraction of a convolutional neural network in
+  our embedding of APL.\@
+
+%  Most practical programming languages of today prioriterise
+%  performance and usability over safety.  As a result, many
+%  useful correctness invariants cannot be expressed in the given
+%  language, typically due to weakness of the chosen type system.
+%  We are looking for non-invasive verification: how to
+%  guarantee a property about the given program without any modifications
+%  to the language toolchain.
+%
+%  We study an approach when a programming language is shallowly
+%  embedded into a theorem prover, and dependent types are used
+%  to encode the properties of interest.   Bringing the embedded
+%  langauge and the properties back into the original language
+%  is the main focus of this work.  We propose a generic extraction
+%  mechanism that we implement in Agda that is based on the concept
+%  of reflection.  Making it possible to translate any shallowly
+%  embedded language in the target backend without the necessity
+%  to modify the theorem prover.
+%  We demonstrate our framework in action by encoding three languages:
+%  Kaleidoskope, SaC and a subset of APL, culminating in the implementation
+%  and extraction of a convolutional neural network.
+\end{abstract}
+\maketitle
+
 \section{Introduction}
@@
 
+% Agda Basics and Extraction
+\input{background}
+
+% PostScript Language and its embedding
+\section{PostScript and its embedding in Agda}
+Demonstrate a few examples.
+
+\section{Extraction}
+\subsection{Framework Overview}
+Explain very briefly that there is a common extraction part that we can (and do)
+abstract away; and there is a language part that we are focusing on.  This can be
+moved towards the end.
+
+\subsection{PostScript Programs Repesentation}
+Explain the AST and show encoding of an example.
+
+\subsection{Normalisation} \label{sec:normalisation}
+It is useful for inlining functions, and creating macros that can use arbitrary expressions
+as long as their applications evaluate to postscript operators.
+This is not essential and can be moved towards the end.
+
+%\subsubsection{Telescopes}
+\subsection{Controlling Reduction} \label{sec:controlling-reduction}
+This is essential, as we need to avoid inlining of the base operators, otherwsie
+we won't be able to extact our pograms one-to-one.  Explain the mechanism.  Say that we have
+added this to Agda?
+
+\subsection{\label{sec:maptypes}Analysing Types}
+Explain that all we need to do is to find the position of the stack argument, as well as
+verify that the return type is also a stack argument.  In between we can have arbitrary
+number of computationally irrelevant arguments.
+
+\subsection{Pattern Matching}
+The natural way to express functions in Agda is by means of pattern-matching, so it makes sense
+to support it.  All we allow in patternmatching is to ``look'' at the stack argument and the
+individual elements.  But we are not allowed to reshuffle the elements, we have to pass the
+pattern exactly as it was matched (in some sense it is read-only pattern matching).
+
+\subsection{Translating terms} \label{sec:translating-terms}
+The core of the extraction is how do we translate terms.  Also this is the simplest part in
+case of the PostScript example.
+
+\subsection{Example}
+Explain the entire extaction process for the given example?
+
+Shall we talk about termination and how do we deal with that?
+
+\subsection{\label{sec:rewriting}Rewriting}
+We can mention that rewriting is super useful feature that can be used as an optimisation
+prior to extraction.
+
+\section{Scaling to Bigger Langauges}
+Not sure whether this deserves a separate section, but we need to explain
+that the proposed technology perfectly scales to the level of bigger (more complex)
+languages.
+
+\section{Related Work}
+
+\section{Conclusios and Future Work}
+
+%% Bibliography
+\bibliography{paper}
+
 \end{document}

### 2021-07-01 paper.tex
Drafting new abstract.

@@
 %% Title information
-\title[]{Extraction once again}
+\title[]{The Power of Dependent Reflection}
 
@@
 \begin{abstract}
-  Dependently-typed host languages empower users to verify a wide
-  range of properties of embedded languages and
-  programs written in them.  Designers of such embedded languages
-  are faced with a difficult choice between using a shallow or a deep
-  embedding.  The former is easier to use because the entire
-  infrastructure of the host langauge is immediately
-  available. Meanwhile, the latter gives full access to the
-  structure of embedded programs, but is difficult to use in practice,
-  especially when the embedded language is itself dependently typed.
 
-  The main insight presented in this paper is that the choice between
-  shallow and deep embedding can be eliminated by working in a host
-  language with reflection capabilities: we start from a shallow
-  embedding that can use all libraries and tools of the host language,
-  and later use reflection to expose the deep structure of the
-  embedded programs.
+  In this paper we explore a way to write correct by construction
+  programs in languages that do not have built-in verification
+  capabilities.  This is achieved in three steps.  Firstly, we
+  shallowly embed the language (or its subset) inside of a
+  dependently-typed system.  Secondly, we write a program
+  within that embedding using dependent types to guarantee
+  any property of interest.  Thirdly, we provide a way to
+  translate the embedded program back into the original language,
+  \eg{} so that it can be used by the existing compilers.
+
+  The main focus of the paper is the third step, which we
+  call \emph{extraction}.  The main insight of the paper is
+  the use of reflection mechanism of a dependently-typed language
+  to achieve extraction.  As a result, the embedding, the
+  program and the extractor are written in a single language.
+
+  To demonstrate these ideas precisely we embed a subset of
+  the PostScript language in Agda and explain all the three
+  steps.
+
+%  Dependently-typed host languages empower users to verify a wide
+%  range of properties of embedded languages and
+%  programs written in them.  Designers of such embedded languages
+%  are faced with a difficult choice between using a shallow or a deep
+%  embedding.  The former is easier to use because the entire
+%  infrastructure of the host langauge is immediately
+%  available. Meanwhile, the latter gives full access to the
+%  structure of embedded programs, but is difficult to use in practice,
+%  especially when the embedded language is itself dependently typed.
 %
-  Concretely, we apply this technique to embed three
-  programming languages --- Kaleidoscope, SaC, and (a subset of) APL ---
-  into the dependently typed theorem prover Agda, using dependent
-  types to statically enforce several properties of interest.  We then
-  use Agda's reflection capabilities to extract the embedded programs
-  back into the original language, so that the existing toolchain can
-  be leveraged.
-  %
-  In this process, statically verified properties of the host language
-  are mapped onto runtime checks in the target language, allowing
-  extracted programs to interact safely with existing code.
-  %
-  Finally, we demonstrate the feasibility of our approach with the
-  implementation and extraction of a convolutional neural network in
-  our embedding of APL.\@
+%  The main insight presented in this paper is that the choice between
+%  shallow and deep embedding can be eliminated by working in a host
+%  language with reflection capabilities: we start from a shallow
+%  embedding that can use all libraries and tools of the host language,
+%  and later use reflection to expose the deep structure of the
+%  embedded programs.
+%  %
+%  Concretely, we apply this technique to embed three
+%  programming languages --- Kaleidoscope, SaC, and (a subset of) APL ---
+%  into the dependently typed theorem prover Agda, using dependent
+%  types to statically enforce several properties of interest.  We then
+%  use Agda's reflection capabilities to extract the embedded programs
+%  back into the original language, so that the existing toolchain can
+%  be leveraged.
+%  %
+%  In this process, statically verified properties of the host language
+%  are mapped onto runtime checks in the target language, allowing
+%  extracted programs to interact safely with existing code.
+%  %
+%  Finally, we demonstrate the feasibility of our approach with the
+%  implementation and extraction of a convolutional neural network in
+%  our embedding of APL.\@
 

### 2021-07-02 paper.tex
First attempt at introduction.

@@
   within that embedding using dependent types to guarantee
-  any property of interest.  Thirdly, we provide a way to
+  a property of interest.  Thirdly, we provide a way to
   translate the embedded program back into the original language,
@@
 \section{Introduction}
-Right now we have the following plan about the structure of the GPCE paper.
-\begin{enumerate}
-    \item Goal: implementing DSL with dependent type system
-    \item Problem: current two approaches are either doing a deep embedding or
-        implementing a typechecker from scratch
-    \item Reflection to the rescue: we can use reflection for implementing
-        custom compilers
-    \item Our approach by example: demonstrate that this actually works in Agda
-        (ps example)
-    \item Extracting definitions by pattern matching: matching on the stack
-    \item Using normalization to get rid of meta-functions (by using this
-        technique, you get macros for free! / or use rewrite rules example:
-        exchange o exchange = id)
-    \item General technology for writing extractors
-    \item This actually scales to realistic and useful examples!
-\end{enumerate}
 
-Optional parts:
-\begin{enumerate}
-    \item Generating assertions
-    \item Ensuring termination of recursive functions
-\end{enumerate}
+It is often desirable to guarantee that a certain class of errors does
+not occur in the given program.  For example, the program can never dereference
+a null pointer; or division by zero never occurs; or indexing is always
+within bounds.  In practical programming languages, the kind of guarantees
+that we can provide formally is typically very limited.  At the same time,
+there could be many arguments in favour of keeping using these languages.
+For example they may come with great tooling, compilers and libraries.
+Therefore, the problem we are investigating in this paper is how to
+provide formal guarantees for the program in some language, yet being
+able to use the original compilers and tools.
+
+One way to tackle this problem is via rephrasing it in terms
+of dependent types.  If the language of interest were to support dependent
+types, we could encode any property about the given program in
+types.  Then, if the program typechecks, the property holds automatically.
+So, the question now becomes how can we add dependent types to a language
+(for verification part), and then eliminate them (for using the original
+compiler).
+
+While implementing a custom typechecker from scratch for a
+dependently-typed language is possible, it is time-consuming and error-prone.
+Alternatively, one can reuse existing implementations of dependently-type
+systems by means of embedding.  In this case, we can share the type system
+of the host language to typecheck the embedded language.
+
+There are two common approaches to language embedding: deep and shallow.
+Shallow embedding defines a language in terms of functions and operators
+of the host language.  Programs in a shallowly-embedded language are programs
+in the host language, so they are immediately runnable; types of the embedded
+language are types of the host language, so the host typechecker is used for
+checking embedded programs.  The downside of this approach is that we cannot
+easily get the representation of the embedded program.
+
+Deep embedding defines a data type to represent the AST of the embedded language.
+Then we need to define a type system and evaluator for the language separately.
+With a dependently-typed host language the AST and the type system can be
+defined within a single data structure, which is typically referred as
+tagless interpreters~\cite{}.  The benefit of deep embedding is that we
+get immediate access to the structure of the embedded program.
+The approach works beautifully for
+embedded languages with simple type systems.  If the language we want to
+embed is dependently-typed, it is shown~\cite{} to be possible, but the
+resulting embedding is very difficult to use in practice.
+
+Our solution to this problem is to rely on a reflection mechanism that
+can be found in dependently-typed systems such as Idris, Agda, Lean.
+Reflection makes it possible to obtain a representation of the
+given term of the host language.  Essentially, shallow embedding in
+a host language with reflections suddenly becomes deep.
+
+In this paper we demonstrate that the proposed approach is actually possible
+by example.  We embed a subset of PostScript~\cite{}, which is a stack-based
+language used in the corresponding document format.  Simplicity of the language
+allows us to focus on the essence of the proposed approach and avoid unrelated
+technical details.  We use Agda as our host language, and we demonstrate the
+embedding, how to specify a program, and most importantly, program extraction.
+
+\todo[inline]{Should we talk about using rewrites as an optimisation?
+        Should we talk here about scaling to realistic examples?
+        Should we mention our work on extending Agda?}
+\todo[inline]{Technical contributions}
+
+% Right now we have the following plan about the structure of the GPCE paper.
+% \begin{enumerate}
+%     \item Goal: implementing DSL with dependent type system
+%     \item Problem: current two approaches are either doing a deep embedding or
+%         implementing a typechecker from scratch
+%     \item Reflection to the rescue: we can use reflection for implementing
+%         custom compilers
+%     \item Our approach by example: demonstrate that this actually works in Agda
+%         (ps example)
+%     \item Extracting definitions by pattern matching: matching on the stack
+%     \item Using normalization to get rid of meta-functions (by using this
+%         technique, you get macros for free! / or use rewrite rules example:
+%         exchange o exchange = id)
+%     \item General technology for writing extractors
+%     \item This actually scales to realistic and useful examples!
+% \end{enumerate}
+%
+% Optional parts:
+% \begin{enumerate}
+%     \item Generating assertions
+%     \item Ensuring termination of recursive functions
+% \end{enumerate}
 
 % Agda Basics and Extraction
-\input{background}
+\input{latex/background}
 
@@
 
-\subsection{Normalisation} \label{sec:normalisation}
+\subsection{\label{sec:normalisation}Normalisation}
 It is useful for inlining functions, and creating macros that can use arbitrary expressions
@@
 %\subsubsection{Telescopes}
-\subsection{Controlling Reduction} \label{sec:controlling-reduction}
+\subsection{\label{sec:controlling-reduction}Controlling Reduction}
 This is essential, as we need to avoid inlining of the base operators, otherwsie
@@
 
-\subsection{Translating terms} \label{sec:translating-terms}
+\subsection{\label{sec:translating-terms}Translating terms}
 The core of the extraction is how do we translate terms.  Also this is the simplest part in

### 2021-07-05 paper.tex
Introduce PostScript (section 2).
Abstact polishing [take 1]

@@
 \usepackage[all,british]{foreign}
+
+\usepackage{epsf}
 
@@
 \begin{abstract}
-
-  In this paper we explore a way to write correct by construction
+  This paper proposes a novel tecnhique to write correct by construction
   programs in languages that do not have built-in verification
   capabilities.  This is achieved in three steps.  Firstly, we
-  shallowly embed the language (or its subset) inside of a
+  shallowly embed the language (or its subset) into a
   dependently-typed system.  Secondly, we write a program
-  within that embedding using dependent types to guarantee
-  a property of interest.  Thirdly, we provide a way to
+  in that embedding, and we use dependent types to guarantee
+  correctness properties of interest.  Thirdly, we
   translate the embedded program back into the original language,
-  \eg{} so that it can be used by the existing compilers.
+  so that it can be used by the existing compilers and tools.
 
-  The main focus of the paper is the third step, which we
-  call \emph{extraction}.  The main insight of the paper is
-  the use of reflection mechanism of a dependently-typed language
-  to achieve extraction.  As a result, the embedding, the
-  program and the extractor are written in a single language.
+  The main insight of the paper is the observation that de\-pen\-den\-tly-typed
+  languages that are equipped with reflection mecahnisms make it possible
+  to expres all the three steps in a single language.  Essentially,
+  this allows us to express programs, its formal properties and
+  compilers hand-in-hand, offering a lot of flexibility to programmers.
 
-  To demonstrate these ideas precisely we embed a subset of
-  the PostScript language in Agda and explain all the three
-  steps.
+  To demonstrate the approach precisely, we embed a subset of
+  the PostScript language in Agda.  We explain all the three
+  steps, illustrating them with several short examples.
 
@@
 \input{latex/background}
-
-% PostScript Language and its embedding
-\section{PostScript and its embedding in Agda}
-Demonstrate a few examples.
+\input{latex/psembedding}
 

### 2021-07-05 paper.tex
Some additions and changes to the abstract

@@
 \begin{abstract}
-  This paper proposes a novel tecnhique to write correct by construction
-  programs in languages that do not have built-in verification
+
+  Dependent types can be used to ensure programs are correct by construction,
+  but they are not available in most
+  languages and adding them to an existing language is a difficult task.
+  This paper proposes a novel technique to write correct-by-construction
+  programs in languages without built-in verification
   capabilities.  This is achieved in three steps.  Firstly, we
-  shallowly embed the language (or its subset) into a
-  dependently-typed system.  Secondly, we write a program
+  give a shallow embedding of the language (or a subset) into a
+  dependently typed language.  Secondly, we write a program
   in that embedding, and we use dependent types to guarantee
@@
 
-  The main insight of the paper is the observation that de\-pen\-den\-tly-typed
-  languages that are equipped with reflection mecahnisms make it possible
-  to expres all the three steps in a single language.  Essentially,
-  this allows us to express programs, its formal properties and
-  compilers hand-in-hand, offering a lot of flexibility to programmers.
-
-  To demonstrate the approach precisely, we embed a subset of
-  the PostScript language in Agda.  We explain all the three
-  steps, illustrating them with several short examples.
+  Our main insight is that it is possible to
+  express all three steps in a single language, as long as it supports
+  both dependent types and reflection. Essentially,
+  this allows us to express a program, its formal properties, and
+  a compiler for it hand-in-hand, offering a lot of flexibility to programmers.
+  We demonstrate this three-step approach by embedding a subset of
+  the PostScript language in Agda, and illustrating it with several short examples.
+  Thus we use the power of reflection to bring the benefits of dependent types to languages
+  that had to go without them so far.
 

### 2021-07-05 paper.tex
Abstract polishing [take 2]

@@
 
-  Dependent types can be used to ensure programs are correct by construction,
-  but they are not available in most
-  languages and adding them to an existing language is a difficult task.
-  This paper proposes a novel technique to write correct-by-construction
-  programs in languages without built-in verification
-  capabilities.  This is achieved in three steps.  Firstly, we
+  Most existing programming languages provde very little support to formally
+  state and prove properties about programs.  Adding such capabilities is far
+  from trivial, as it requires significant reengineering of the existing
+  compilers and tools.  This paper proposes a novel technique to write
+  correct-by-construction programs in languages without built-in verification
+  capabilities, maintaining the ability to use existing tools.
+  This is achieved in three steps.  Firstly, we
   give a shallow embedding of the language (or a subset) into a

### 2021-07-05 paper.tex
Abstract typo + nitpicking

@@
 
-  Most existing programming languages provde very little support to formally
+  Most existing programming languages provide little support to formally
   state and prove properties about programs.  Adding such capabilities is far
-  from trivial, as it requires significant reengineering of the existing
+  from trivial, as it requires significant re-engineering of the existing
   compilers and tools.  This paper proposes a novel technique to write

### 2021-07-05 paper.tex
Add code for main extraction example

@@
 \usepackage{varwidth}
+\usepackage{microtype}
 
@@
 \usepackage[final]{listings}
+
+\usepackage[T1]{fontenc}
 
@@
 \newunicodechar{⌿}{\ensuremath{\not-}}
+
+\newunicodechar{←}{\ensuremath{\leftarrow}}
+\newunicodechar{×}{\ensuremath{\times}}
+\newunicodechar{λ}{\ensuremath{\lambda}}
+\newunicodechar{∀}{\ensuremath{\forall}}
+\newunicodechar{Π}{\ensuremath{\Pi}}
+\newunicodechar{≡}{\ensuremath{\equiv}}
+\newunicodechar{≅}{\ensuremath{\cong}}
+\newunicodechar{≐}{\ensuremath{\doteq}}
+\newunicodechar{∈}{\ensuremath{\in}}
+\newunicodechar{∧}{\ensuremath{\land}}
+\newunicodechar{∨}{\ensuremath{\lor}}
+\newunicodechar{⊤}{\ensuremath{\top}}
+\newunicodechar{⊥}{\ensuremath{\bot}}
+\newunicodechar{⊔}{\ensuremath{\sqcup}}
+\newunicodechar{∷}{\ensuremath{{::}}}
+\newunicodechar{ℓ}{\ensuremath{\ell}}
+\newunicodechar{₀}{\ensuremath{{_0}}}
+\newunicodechar{₁}{\ensuremath{{_1}}}
+\newunicodechar{₂}{\ensuremath{{_2}}}
+\newunicodechar{₃}{\ensuremath{{_3}}}
+\newunicodechar{₄}{\ensuremath{{_4}}}
+\newunicodechar{₅}{\ensuremath{{_5}}}
+\newunicodechar{₆}{\ensuremath{{_6}}}
+\newunicodechar{₇}{\ensuremath{{_7}}}
+\newunicodechar{₈}{\ensuremath{{_8}}}
+\newunicodechar{₉}{\ensuremath{{_9}}}
+\newunicodechar{⟨}{\ensuremath{{\langle}}}
+\newunicodechar{⟩}{\ensuremath{{\rangle}}}
+\newunicodechar{̧}{\c}
+\newunicodechar{≤}{\ensuremath{\leq}}
+\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}
+\newunicodechar{∸}{̣\ensuremath{-}}
+\newunicodechar{ᵇ}{\ensuremath{{}^{b}}}
 
@@
 \input{latex/psembedding}
-
-\section{Extraction}
-\subsection{Framework Overview}
-Explain very briefly that there is a common extraction part that we can (and do)
-abstract away; and there is a language part that we are focusing on.  This can be
-moved towards the end.
-
-\subsection{PostScript Programs Repesentation}
-Explain the AST and show encoding of an example.
-
-\subsection{\label{sec:normalisation}Normalisation}
-It is useful for inlining functions, and creating macros that can use arbitrary expressions
-as long as their applications evaluate to postscript operators.
-This is not essential and can be moved towards the end.
-
-%\subsubsection{Telescopes}
-\subsection{\label{sec:controlling-reduction}Controlling Reduction}
-This is essential, as we need to avoid inlining of the base operators, otherwsie
-we won't be able to extact our pograms one-to-one.  Explain the mechanism.  Say that we have
-added this to Agda?
-
-\subsection{\label{sec:maptypes}Analysing Types}
-Explain that all we need to do is to find the position of the stack argument, as well as
-verify that the return type is also a stack argument.  In between we can have arbitrary
-number of computationally irrelevant arguments.
-
-\subsection{Pattern Matching}
-The natural way to express functions in Agda is by means of pattern-matching, so it makes sense
-to support it.  All we allow in patternmatching is to ``look'' at the stack argument and the
-individual elements.  But we are not allowed to reshuffle the elements, we have to pass the
-pattern exactly as it was matched (in some sense it is read-only pattern matching).
-
-\subsection{\label{sec:translating-terms}Translating terms}
-The core of the extraction is how do we translate terms.  Also this is the simplest part in
-case of the PostScript example.
-
-\subsection{Example}
-Explain the entire extaction process for the given example?
-
-Shall we talk about termination and how do we deal with that?
-
-\subsection{\label{sec:rewriting}Rewriting}
-We can mention that rewriting is super useful feature that can be used as an optimisation
-prior to extraction.
+\input{latex/extraction}
 

### 2021-07-08 paper.tex
Title
Commenting out fontenc
First version of section 5
Copy-paste related work and conclusion from ICFP submission (with some things removed)

@@
 
-\usepackage[T1]{fontenc}
+%\usepackage[T1]{fontenc}
 
@@
 %% Title information
-\title[]{The Power of Dependent Reflection}
+\title[]{Extracting The Power of Dependent Types}
 
@@
 \input{latex/extraction}
+\input{latex/reduction}
 
@@
 
-\section{Related Work}
+\section{\label{sec:related}Related Work}
 
-\section{Conclusios and Future Work}
+\paragraph{Metaprogramming} There is a large body of
+work on metaprogramming facilities in various programming
+languages.  \citet{refl-masses} track the origins
+of metaprogramming to Smith's work on reflection in Lisp~\cite{refl-lisp}.
+Some prominent metaprogramming systems include
+MetaOcaml~\cite{metaocaml}, MetaML~\cite{metaml},
+reFlect~\cite{DBLP:journals/jfp/GrundyMO06},
+Template Haskell~\cite{sheard2002template},
+Racket~\cite{plt-tr1}, and various other Lisp/Scheme dialects.
+%
+However, these systems typically do not support dependent
+types, so they are not well suited for our goal of statically
+enforcing correctness of embedded programs.
+
+\paragraph{Embedding}
+Defining deep embeddings with static guarantees are a common application
+of dependent types~\cite{10.5555/647849.737066,CHAPMAN200921,
+10.1007/978-3-540-74464-1_7,10.1145/3236785,10.1145/1863495.1863497}.
+%
+These embeddings usually also define semantics of the embedded
+language and therefore allow us to reason about the correctness
+of program transformations and optimisations.
+%
+While the fact that this is possible is impressive in theory, the resulting
+encodings are often difficult to use in practice. In this paper
+we instead aim for a more lightweight approach.
+
+\citet{deepshallow} propose to solve this problem with
+a combination of deep and shallow embeddings.  Their idea is to define
+a small deep embedding and leverage type classes in Haskell to define
+the rest of the language on top of that.  It would be interesting to see
+whether such an approach scales to dependently-typed embedded
+languages.
+
+\paragraph{Extraction}
+The Coq proof assistant is equipped with extraction
+capabilities~\cite{10.1007/978-3-540-69407-6_39,10.1007/3-540-39185-1_12},
+which extracts functional code from Coq proofs (or programs).  The
+default target language is Ocaml, but a few other options were added
+recently.
+%
+Likewise, Agda itself has a mechanism for defining custom backends, of
+which the GHC backend is the most prominent.
+%
+Other proof assistants provide similar extraction tools as well.
+%
+The main difference from our approach in this paper
+is that these extractors are written as plugins to the proof
+assistant, while we implement our extractors directly in the proof
+assistant itself.
+%
+While it would be possible to implement extractors presented in this
+paper as Coq plugins or Agda backends, conceptually they are more
+heavyweight.  Our extractors and programs can (in principle)
+communicate with each other. In addition, as they are just Agda programs, they can
+be reflected themselves and their structure can be leveraged.
+
+\paragraph{Dependently typed metaprogramming}
+Several dependently-typed languages are equipped with metaprogramming
+capabilities: Idris \cite{idris-refl}, Lean \cite{lean-refl},
+Coq \cite{metacoq}, and Agda \cite{agda-refl}.  All of these
+implement a similar API as described in this paper.  This is
+reassuring, as it means our proposed approach is immediately portable
+into many other contexts.
+\citet{10.1145/3371071} introduce the Turnstile+ framework for
+building dependently typed embedded DSLs and very much
+shares the ideas advocated in this paper, suggesting that
+our approach could work there as well.
+\citet{10.1145/3371076} use MetaCoq to formally verify
+the core type system of Coq. This combines nicely with
+our approach, as we could use the verified core language
+as a basis to verify our custom extractors.
+\citet{10.1145/3372885.3373829} use MetaCoq to implement
+a DSL combining deep and shallow approaches, in a way that
+is quite similar to our own. While they are able to formally
+reason about preservation of semantics (which we can't do yet), it is unclear
+whether their approach scales to dependently-typed embedded languages.
+
+\section{Conclusions and Future Work}
+
+In this paper we investigate the idea of developing
+embedded programs hand-in-hand with custom
+code generators for them. We solve the well-known
+conundrum of choosing between deep and shallow
+embedding by leveraging the power of
+reflection.
+This allows us to enjoying the benefits
+of shallow embedding, while keeping full access to the internal
+structure of the embedded programs.
+
+We apply this idea in the context of dependently-typed
+languages to create verified implementations that can
+be used in the context of an existing programming language.
+We embed the target language in a theorem-prover, using
+dependent types to encode the properties of interest.
+Using reflection we bring the verified implementation
+back into the original language.
+%
+Concretely, we have demonstrated the approach by implementing
+an extractor for a fragment of the PostScript language, using
+Agda as our host.
+
+The main advantages of our approach are twofold.  First,
+our solution is fine-grained --- we can chose what part
+of the application to embed, and what constructs of the
+host language to extract.  Secondly, our extractors
+can use the full power of dependent types to guarantee
+safety of our embedded programs.
+
+Right now we cannot yet guarantee that the
+extracted code preservers the semantics of the original
+implementation. While we rarely see
+fully-verified compiler backends in the real world,
+our approach is very close to enabling this.
+We would need a formal semantics of the reflected language
+and the proof that reflected programs respect it.
+While this is non-trivial, a system like Agda could do
+this in principle.
+
+There is a number of improvements that can be added to
+Agda and our extractors to make the resulting code more
+efficient.  Supporting \AK{let}s in the internal syntax
+would help to preserve sharing.  Recognising irrelevance
+annotations in the extractors would help to eliminate
+unused function arguments.  Introducing proper language
+primitives to specify what exactly is an embedding would
+be helpful. And finally, having access to more of
+Agda's internals such as case trees would help to
+generate more performant code.
+
+Overall, this work only scratches the surface of extraction-based
+compilation.  We never considered alternative theories supported
+by theorem provers, \eg{} cubical type theory in Agda; we did not
+consider recursive metaprogramming; we did not consider integrating
+optimisations of extracted programs other than what rewriting rules
+are capable to do.  All of these offer exciting research opportunities
+on the way to make verified software easily accessible.
+To paraphrase Jim Morrison: ``Safety and no surprise, The End''.
 

### 2021-07-09 paper.tex
Pass over abstract + introduction

@@
 
+\newcommand{\myref}[2]{\hyperref[#2]{#1~\ref*{#2}}}
+\newcommand{\figref}[1]{\myref{Fig.}{fig:#1}}%{Fig.~\ref{fig:#1}}
+\newcommand{\secref}[1]{\myref{Sect.}{sec:#1}}
+\newcommand{\lemref}[1]{\myref{Lemma}{lem:#1}}
+\newcommand{\thmref}[1]{\myref{Theorem}{thm:#1}}
+\newcommand{\defref}[1]{\myref{Definition}{def:#1}}
+\newcommand{\eqnref}[1]{\hyperref[#1]{(\ref{#1})}}
+
 \begin{document}
@@
   correct-by-construction programs in languages without built-in verification
-  capabilities, maintaining the ability to use existing tools.
+  capabilities, while maintaining the ability to use existing tools.
   This is achieved in three steps.  Firstly, we
@@
   correctness properties of interest.  Thirdly, we
-  translate the embedded program back into the original language,
+  extract a program written in the original language,
   so that it can be used by the existing compilers and tools.
@@
   Our main insight is that it is possible to
-  express all three steps in a single language, as long as it supports
+  express all three steps in a single language that supports
   both dependent types and reflection. Essentially,
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 
-It is often desirable to guarantee that a certain class of errors does
-not occur in the given program.  For example, the program can never dereference
-a null pointer; or division by zero never occurs; or indexing is always
-within bounds.  In practical programming languages, the kind of guarantees
-that we can provide formally is typically very limited.  At the same time,
-there could be many arguments in favour of keeping using these languages.
-For example they may come with great tooling, compilers and libraries.
-Therefore, the problem we are investigating in this paper is how to
-provide formal guarantees for the program in some language, yet being
-able to use the original compilers and tools.
+When writing a program, it is often desirable to guarantee that a
+certain class of errors does not occur in the given program.  For
+example, the program can never dereference a null pointer; or division
+by zero never occurs; or indexing is always within bounds.  In most
+programming languages, the class of guarantees that can be checked
+statically is very limited.  Nevertheless, moving to a different
+language is not always an option.  For example, these languages may
+come with great tooling, compilers and libraries.  In this paper, we
+investigate how to provide strong static guarantees for programs in a
+given language, while maintaining the ability to to use existing
+compilers and tools.
 
-One way to tackle this problem is via rephrasing it in terms
-of dependent types.  If the language of interest were to support dependent
-types, we could encode any property about the given program in
-types.  Then, if the program typechecks, the property holds automatically.
-So, the question now becomes how can we add dependent types to a language
-(for verification part), and then eliminate them (for using the original
-compiler).
+One very powerful method for enforcing almost any functional property
+of a program statically is by using \emph{dependent types}.
+In a dependently typed language, properties can be encoded as types,
+and if the program typechecks, the property is guaranteed to hold.
+If we can extend our existing language with dependent types
+(for verification part) and then eliminate them (for using the original
+compiler), then our goal is achieved.
 
-While implementing a custom typechecker from scratch for a
-dependently-typed language is possible, it is time-consuming and error-prone.
-Alternatively, one can reuse existing implementations of dependently-type
-systems by means of embedding.  In this case, we can share the type system
+While it is possible to add dependent types to an existing language
+from scratch, doing so is time-consuming and error-prone.  Instead,
+one can also embed the language we want to use into an existing
+dependently typed language, which allows us to re-use the typechecker
 of the host language to typecheck the embedded language.
@@
 There are two common approaches to language embedding: deep and shallow.
-Shallow embedding defines a language in terms of functions and operators
+A shallow embedding defines a language in terms of functions and operators
 of the host language.  Programs in a shallowly-embedded language are programs
@@
 
-Deep embedding defines a data type to represent the AST of the embedded language.
-Then we need to define a type system and evaluator for the language separately.
-With a dependently-typed host language the AST and the type system can be
-defined within a single data structure, which is typically referred as
-tagless interpreters~\cite{}.  The benefit of deep embedding is that we
-get immediate access to the structure of the embedded program.
-The approach works beautifully for
-embedded languages with simple type systems.  If the language we want to
-embed is dependently-typed, it is shown~\cite{} to be possible, but the
-resulting embedding is very difficult to use in practice.
+Meanwhile, a deep embedding defines a data type to represent the AST
+of the embedded language.  This requires us to define a type system
+and evaluator for the language separately.  With a dependently typed
+host language the AST and the type system can be defined as a single
+data structure, which is typically referred to as a tagless
+interpreter~\cite{}.  The benefit of deep embedding is that we get
+full access to the syntactic structure of the embedded program.  This
+approach works beautifully for embedded languages with simple type
+systems. However, for embedded languages with dependent types,
+defining a tagless interpreter is technically possible~\cite{}, but
+extremely challenging to define and use in practice.
 
-Our solution to this problem is to rely on a reflection mechanism that
-can be found in dependently-typed systems such as Idris, Agda, Lean.
-Reflection makes it possible to obtain a representation of the
-given term of the host language.  Essentially, shallow embedding in
-a host language with reflections suddenly becomes deep.
+We propose a novel approach that combines the best parts of shallow
+and deep embeddings through the use of reflection in a
+dependently-typed host language such as Coq, Agda, Idris, or Lean.
+Concretely, we start by defining a shallow embedding of the target
+language while making use of dependent types to statically enforce the
+properties of interest, and write a program in this embedding. We then
+make use of reflection to obtain a syntactic representation of the
+program, and extract a program in the syntax of the target language
+from it.  Finally, we compile the extracted program using existing
+tools and libraries. Essentially, a shallow embedding in a host
+language with reflections suddenly becomes deep.
 
@@
 by example.  We embed a subset of PostScript~\cite{}, which is a stack-based
-language used in the corresponding document format.  Simplicity of the language
+language used in the corresponding document format.  The simplicity of the language
 allows us to focus on the essence of the proposed approach and avoid unrelated
@@
 
-\todo[inline]{Should we talk about using rewrites as an optimisation?
-        Should we talk here about scaling to realistic examples?
-        Should we mention our work on extending Agda?}
-\todo[inline]{Technical contributions}
+\paragraph{Contributions.}
+\begin{itemize}
+
+\item We propose a novel approach that can be used to statically
+  enforce almost any property of programs in an existing language, by
+  embedding it in a host language that supports dependent types and
+  reflection.
+
+\item We demonstate this approach by a concrete example: an embedding
+  of (a subset of the) PostScript language into Agda.
+
+\item We give several examples of PostScript programs written in our
+  embedding, and demonstrate how to statically enforce properties such
+  as the absence of stack underflows, termination, and functional
+  correctness.
+
+\item We show how to use reflection in Agda to implement an extractor
+  that translates code written in our embedding to plain PostScript.
+
+\item We show how to our approach provides easy support for partial
+  evaluation by making use of normalisation in the host language. This
+  allows us to use functions of the host language as macros, as well
+  as automatically apply optimizations through the use of rewrite
+  rules.
+\end{itemize}
+
+The rest of this paper is structured as follows. \secref{background}
+provides background information on Agda and its reflection
+capabilities. \secref{embedding} describes our embedding of PostScript
+in Agda, and gives several examples of programs written in
+it. \secref{extraction} shows how to implement an extractor for this
+embedding using reflection. Next, \secref{partial-evaluation}
+demonstrates how to apply partial evaluation to programs prior to
+extraction by using Agda's normalisation procedure. Finally,
+\secref{related} discusses related work and \secref{conclusion}
+concludes.
 
@@
 
-\section{Scaling to Bigger Langauges}
+\section{Scaling to Bigger Languages} \label{sec:scaling}
 Not sure whether this deserves a separate section, but we need to explain
@@
 
-\section{Conclusions and Future Work}
+\section{Conclusions and Future Work} \label{sec:conclusion}
 

### 2021-07-09 paper.tex
Add paragraph on limitations to introduction

@@
 
+\paragraph{Limitations}
+
+While we claim our approach is flexible and powerful, it is
+certainly not a magic bullet. There are still a lot of things that
+need to be implemented for each target language: the language
+embedding itself, the encoding of the properties under consideration
+as types in the host language, and the extractor that transforms
+embedded programs back into the syntax of the original language.  For
+enforcing more complex properties, it might also be necessary to write
+proofs to convince the typechecker of the host language that they are
+indeed satisfied.
+%
+The main advantage of our approach is twofold: first, it allows
+you to implement all these things --- as well as the embedded programs
+themselves --- \emph{side by side in a single language}, simplifying
+the development. And secondly, it allows you to reuse the typechecker
+and evaluator of the host language, saving you a lot of work.  In the
+end, our goal is not to showcase a single example but to empower you
+to apply this approach yourself to verify code in any language.
+
 \paragraph{Contributions.}

### 2021-07-11 paper.tex
Adjusting sections 3 and 4.
Add paragraph on other languages we did this for

@@
 
+\todo[inline]{Move it somewhere}
+This picture has been generated by the code that was extracted from our
+Agda embedding.
+
+\begin{center}
+\epsfbox[0 0 64 64]{sierp.ps}
+\end{center}
+
 % Right now we have the following plan about the structure of the GPCE paper.
@@
 \input{latex/reduction}
-
-\section{Scaling to Bigger Languages} \label{sec:scaling}
-Not sure whether this deserves a separate section, but we need to explain
-that the proposed technology perfectly scales to the level of bigger (more complex)
-languages.
 
@@
 
+\paragraph{Scaling to bigger languages}
+
+The main goal of this paper is to demonstrate how to define a
+embedding and extractor using dependent types and reflection on a
+small but complete example. In addition to the PostScript embedding
+discussed in this paper, we applied our approach to several other
+languages:
+\begin{itemize}
+\item Kaleidoscope, a minimalist imperative language used in the
+  documentation of LLVM.
+\item SaC (Single assignment C), a high-performance array language
+  that generates efficient code for various architectures.
+\item APL (Array Programming Language), another array language with
+  heavily overloaded syntax that pioneered the concept of rank- and
+  shape-polymorphic programming.
+\end{itemize}
+In addition, we used the embedding of APL to implement a Convolutional
+Neural Network, and ensure that all the array instructions operate on
+arrays of the correct dimensions. For more details on these examples,
+refer to \citet{DBLP:journals/corr/abs-2105-10819}.
+
+\paragraph{Future work}
+
 Right now we cannot yet guarantee that the

### 2021-07-13 paper.tex
Comment out triangle picture for now
Clarify itnroduction.

@@
 
-When writing a program, it is often desirable to guarantee that a
-certain class of errors does not occur in the given program.  For
-example, the program can never dereference a null pointer; or division
-by zero never occurs; or indexing is always within bounds.  In most
-programming languages, the class of guarantees that can be checked
-statically is very limited.  Nevertheless, moving to a different
+It is often desirable to guarantee that a
+certain class of errors does not occur in the given program.  These
+error classes may include: dereferencing the null pointer, division
+by zero, or out of bound indexing.  In most
+programming languages, the guarantees that can be checked
+statically are very limited.  Nevertheless, moving to a different
 language is not always an option.  For example, these languages may
-come with great tooling, compilers and libraries.  In this paper, we
+come with great tooling, compilers and/or libraries.  In this paper, we
 investigate how to provide strong static guarantees for programs in a
@@
 from scratch, doing so is time-consuming and error-prone.  Instead,
-one can also embed the language we want to use into an existing
-dependently typed language, which allows us to re-use the typechecker
+one can also embed the language we want to use into existing
+dependently typed languages, which allows us to re-use the typechecker
 of the host language to typecheck the embedded language.
@@
 checking embedded programs.  The downside of this approach is that we cannot
-easily get the representation of the embedded program.
+easily get the representation of embedded programs.
 
@@
 systems. However, for embedded languages with dependent types,
-defining a tagless interpreter is technically possible~\cite{}, but
-extremely challenging to define and use in practice.
+tagless interpreters are technically possible~\cite{}, but
+are extremely challenging to define and use in practice.
 
-We propose a novel approach that combines the best parts of shallow
-and deep embeddings through the use of reflection in a
+The key insight of this paper lies in proposing a novel approach that
+combines the best parts of shallow and deep embeddings through the use
+of reflection in a
 dependently-typed host language such as Coq, Agda, Idris, or Lean.
-Concretely, we start by defining a shallow embedding of the target
-language while making use of dependent types to statically enforce the
-properties of interest, and write a program in this embedding. We then
-make use of reflection to obtain a syntactic representation of the
-program, and extract a program in the syntax of the target language
-from it.  Finally, we compile the extracted program using existing
-tools and libraries. Essentially, a shallow embedding in a host
-language with reflections suddenly becomes deep.
 
-In this paper we demonstrate that the proposed approach is actually possible
-by example.  We embed a subset of PostScript~\cite{}, which is a stack-based
-language used in the corresponding document format.  The simplicity of the language
+Our main objective of the paper is to demonstrate that the
+proposed approach works in practice.  Therefore, we guide the reader
+through the entire process of: i) shallow embedding of the target
+language;  ii) writing programs in this embedding; iii) using
+dependent types to statically enforce the properties of interest; 
+iv) extract a program in the syntax of the target language
+from it.  We run the extracted program using existing
+tools and libraries. Essentially, demonstrating that a
+shallow embedding in a host language with reflections
+suddenly becomes deep.
+
+In order to keep presentation as simple as possible, we have
+chosen a very simple target language to embed --- we use a subset
+of PostScript~\cite{}, which is a stack-based
+language used in the corresponding document format.  While the embedding
+may look impractical, the simplicity of the language
 allows us to focus on the essence of the proposed approach and avoid unrelated
-technical details.  We use Agda as our host language, and we demonstrate the
-embedding, how to specify a program, and most importantly, program extraction.
-
-\paragraph{Limitations}
-
-While we claim our approach is flexible and powerful, it is
-certainly not a magic bullet. There are still a lot of things that
-need to be implemented for each target language: the language
-embedding itself, the encoding of the properties under consideration
-as types in the host language, and the extractor that transforms
-embedded programs back into the syntax of the original language.  For
-enforcing more complex properties, it might also be necessary to write
-proofs to convince the typechecker of the host language that they are
-indeed satisfied.
-%
-The main advantage of our approach is twofold: first, it allows
-you to implement all these things --- as well as the embedded programs
-themselves --- \emph{side by side in a single language}, simplifying
-the development. And secondly, it allows you to reuse the typechecker
-and evaluator of the host language, saving you a lot of work.  In the
-end, our goal is not to showcase a single example but to empower you
-to apply this approach yourself to verify code in any language.
+technical details.  We ensure that many verification patterns found in bigger
+languages are covered, implying that the proposed approach scales.
+We use Agda as our host language, and the entire paper is Agda script.
+This means that the code presented in the snippets has been typechecked
+and presented examples ran.
 
@@
 
-\item We demonstate this approach by a concrete example: an embedding
+\item We demonstrate this approach by a concrete example: an embedding
   of (a subset of the) PostScript language into Agda.
@@
 
+\begin{comment}
 \todo[inline]{Move it somewhere}
@@
 \end{center}
+\end{comment}
 
@@
 
+\paragraph{Limitations}
+\todo[inline]{This has moved from the intro, needs reflowing!}
+While we claim our approach is flexible and powerful, it is
+certainly not a magic bullet. There are still a lot of things that
+need to be implemented for each target language: the language
+embedding itself, the encoding of the properties under consideration
+as types in the host language, and the extractor that transforms
+embedded programs back into the syntax of the original language.  For
+enforcing more complex properties, it might also be necessary to write
+proofs to convince the typechecker of the host language that they are
+indeed satisfied.
+%
+The main advantage of our approach is twofold: first, it allows
+you to implement all these things --- as well as the embedded programs
+themselves --- \emph{side by side in a single language}, simplifying
+the development. And secondly, it allows you to reuse the typechecker
+and evaluator of the host language, saving you a lot of work.  In the
+end, our goal is not to showcase a single example but to empower you
+to apply this approach yourself to verify code in any language.
+
 \paragraph{Scaling to bigger languages}

### 2021-07-13 paper.tex
Brushed conclusions.

@@
 reflection.
-This allows us to enjoying the benefits
-of shallow embedding, while keeping full access to the internal
-structure of the embedded programs.
+This allows us to enjoy simplicity of writing progams
+in shallow embedding.  Yet we keep full access to the internal
+structure of the embedded programs, so that we could translate
+them into the original language.
 
 We apply this idea in the context of dependently-typed
-languages to create verified implementations that can
-be used in the context of an existing programming language.
-We embed the target language in a theorem-prover, using
-dependent types to encode the properties of interest.
-Using reflection we bring the verified implementation
-back into the original language.
-%
+language Agda.  We demonstrate the use of dependent types
+in order to enforce static propeties of the embedded programs.
 Concretely, we have demonstrated the approach by implementing
-an extractor for a fragment of the PostScript language, using
-Agda as our host.
+an extractor for a fragment of the PostScript language.
 
-The main advantages of our approach are twofold.  First,
-our solution is fine-grained --- we can chose what part
-of the application to embed, and what constructs of the
-host language to extract.  Secondly, our extractors
-can use the full power of dependent types to guarantee
-safety of our embedded programs.
-
-\paragraph{Limitations}
-\todo[inline]{This has moved from the intro, needs reflowing!}
-While we claim our approach is flexible and powerful, it is
-certainly not a magic bullet. There are still a lot of things that
-need to be implemented for each target language: the language
-embedding itself, the encoding of the properties under consideration
-as types in the host language, and the extractor that transforms
-embedded programs back into the syntax of the original language.  For
-enforcing more complex properties, it might also be necessary to write
-proofs to convince the typechecker of the host language that they are
-indeed satisfied.
-%
 The main advantage of our approach is twofold: first, it allows
@@
 the development. And secondly, it allows you to reuse the typechecker
-and evaluator of the host language, saving you a lot of work.  In the
-end, our goal is not to showcase a single example but to empower you
-to apply this approach yourself to verify code in any language.
-
-\paragraph{Scaling to bigger languages}
-
-The main goal of this paper is to demonstrate how to define a
-embedding and extractor using dependent types and reflection on a
-small but complete example. In addition to the PostScript embedding
-discussed in this paper, we applied our approach to several other
+and evaluator of the host language, saving you a lot of work.
+While we showcase a single example, we applied our approach to several other
 languages:
@@
 \end{itemize}
-In addition, we used the embedding of APL to implement a Convolutional
-Neural Network, and ensure that all the array instructions operate on
-arrays of the correct dimensions. For more details on these examples,
-refer to \citet{DBLP:journals/corr/abs-2105-10819}.
+For more details on these examples,
+refer to~\citet{DBLP:journals/corr/abs-2105-10819}.
+This sugests that you are now empowered to apply the proposed
+technique tp verify code in any language.
 
 \paragraph{Future work}
+While our approach is flexible and powerful, it is
+certainly not a magic bullet. A lot of things
+need to be implemented for each target language: the embedding,
+the encoding of the properties under consideration as types,
+and the extractor.  For enforcing more complex properties, a number
+of proofs needs to be written explicitly.
 
-Right now we cannot yet guarantee that the
+Right now, we cannot yet guarantee that the
 extracted code preservers the semantics of the original
@@
 
-There is a number of improvements that can be added to
-Agda and our extractors to make the resulting code more
-efficient.  Supporting \AK{let}s in the internal syntax
-would help to preserve sharing.  Recognising irrelevance
-annotations in the extractors would help to eliminate
-unused function arguments.  Introducing proper language
-primitives to specify what exactly is an embedding would
-be helpful. And finally, having access to more of
-Agda's internals such as case trees would help to
-generate more performant code.
-
-Overall, this work only scratches the surface of extraction-based
-compilation.  We never considered alternative theories supported
-by theorem provers, \eg{} cubical type theory in Agda; we did not
-consider recursive metaprogramming; we did not consider integrating
-optimisations of extracted programs other than what rewriting rules
-are capable to do.  All of these offer exciting research opportunities
-on the way to make verified software easily accessible.
-To paraphrase Jim Morrison: ``Safety and no surprise, The End''.
+% There is a number of improvements that can be added to
+% Agda and our extractors to make the resulting code more
+% efficient.  Supporting \AK{let}s in the internal syntax
+% would help to preserve sharing.  Recognising irrelevance
+% annotations in the extractors would help to eliminate
+% unused function arguments.  Introducing proper language
+% primitives to specify what exactly is an embedding would
+% be helpful. And finally, having access to more of
+% Agda's internals such as case trees would help to
+% generate more performant code.
+% 
+% Overall, this work only scratches the surface of extraction-based
+% compilation.  We never considered alternative theories supported
+% by theorem provers, \eg{} cubical type theory in Agda; we did not
+% consider recursive metaprogramming; we did not consider integrating
+% optimisations of extracted programs other than what rewriting rules
+% are capable to do.  All of these offer exciting research opportunities
+% on the way to make verified software easily accessible.
+% To paraphrase Jim Morrison: ``Safety and no surprise, The End''.
 

### 2021-07-13 paper.tex
typos

@@
 reflection.
-This allows us to enjoy simplicity of writing progams
+This allows us to enjoy simplicity of writing programs
 in shallow embedding.  Yet we keep full access to the internal
@@
 language Agda.  We demonstrate the use of dependent types
-in order to enforce static propeties of the embedded programs.
+in order to enforce static properties of the embedded programs.
 Concretely, we have demonstrated the approach by implementing
@@
 refer to~\citet{DBLP:journals/corr/abs-2105-10819}.
-This sugests that you are now empowered to apply the proposed
-technique tp verify code in any language.
+This suggests that you are now empowered to apply the proposed
+technique to verify code in any language.
 

### 2021-07-13 paper.tex
Anonymity + space reduction

@@
 and evaluator of the host language, saving you a lot of work.
-While we showcase a single example, we applied our approach to several other
-languages:
-\begin{itemize}
-\item Kaleidoscope, a minimalist imperative language used in the
-  documentation of LLVM.
-\item SaC (Single assignment C), a high-performance array language
-  that generates efficient code for various architectures.
-\item APL (Array Programming Language), another array language with
-  heavily overloaded syntax that pioneered the concept of rank- and
-  shape-polymorphic programming.
-\end{itemize}
+While we showcase a single example, the approach has been applied
+to several other languages: Kaleidoscope, a minimalist imperative language;
+Single assignment C, a high-performance array language;
+APL (Array Programming Language), another array language with
+heavily overloaded syntax.
 For more details on these examples,
 refer to~\citet{DBLP:journals/corr/abs-2105-10819}.
-This suggests that you are now empowered to apply the proposed
+
+We believe that the reader is now empowered to apply the proposed
 technique to verify code in any language.

### 2021-07-13 paper.tex
Small tweaks

@@
 programming languages, the guarantees that can be checked
-statically are very limited.  Nevertheless, moving to a different
+statically are limited.  Nevertheless, moving to a different
 language is not always an option.  For example, these languages may
@@
 
-One very powerful method for enforcing almost any functional property
+One powerful method for enforcing almost any functional property
 of a program statically is by using \emph{dependent types}.
@@
 from scratch, doing so is time-consuming and error-prone.  Instead,
-one can also embed the language we want to use into existing
-dependently typed languages, which allows us to re-use the typechecker
+we can \emph{embed} the language we want to use into an existing
+dependently typed host language, which allows us to re-use the typechecker
 of the host language to typecheck the embedded language.
@@
 dependent types to statically enforce the properties of interest;
-iv) extract a program in the syntax of the target language
+iv) extracting a program in the syntax of the target language
 from it.  We run the extracted program using existing
-tools and libraries. Essentially, demonstrating that a
-shallow embedding in a host language with reflections
+tools and libraries. Essentially, we demonstrate that a
+shallow embedding in a host language with reflection
 suddenly becomes deep.
@@
 technical details.  We ensure that many verification patterns found in bigger
-languages are covered, implying that the proposed approach scales.
-We use Agda as our host language, and the entire paper is Agda script.
-This means that the code presented in the snippets has been typechecked
-and presented examples ran.
+languages are covered, showing that the proposed approach scales.
+
+This entire paper is a literate Agda script: all Agda code is typechecked
+and all the examples are run while generating the paper.
 
@@
 to several other languages: Kaleidoscope, a minimalist imperative language;
-Single assignment C, a high-performance array language;
+Single assignment C, a high-performance array language; and
 APL (Array Programming Language), another array language with

### 2021-07-14 paper.tex
Move an example to the introduction

@@
 
-In order to keep presentation as simple as possible, we have
+\input{latex/intro-example}
+
+To keep presentation as simple as possible, we have
 chosen a very simple target language to embed --- we use a subset
-of PostScript~\cite{}, which is a stack-based
-language used in the corresponding document format.  While the embedding
-may look impractical, the simplicity of the language
+of PostScript~\citep{postscript1999language}, a stack-based
+language used in the corresponding document format.
+%
+\figref{sqsum} shows an example of a program
+written in our embedding and the PostScript code produced by
+our extractor.
+%
+While the embedding may look simplistic, this simplicity
 allows us to focus on the essence of the proposed approach and avoid unrelated

### 2021-07-14 paper.tex
Add some missing citations

@@
 
-One powerful method for enforcing almost any functional property
-of a program statically is by using \emph{dependent types}.
-In a dependently typed language, properties can be encoded as types,
-and if the program typechecks, the property is guaranteed to hold.
-If we can extend our existing language with dependent types
-(for verification part) and then eliminate them (for using the original
-compiler), then our goal is achieved.
+One powerful method for enforcing almost any functional property of a
+program statically is by using \emph{dependent types}, as used for
+example in Coq~\cite{Coq-8-13-2} or Agda~\cite{agda-2-6-2}.  In a
+dependently typed language, properties can be encoded as types, and if
+the program typechecks, the property is guaranteed to hold.  If we can
+extend our existing language with dependent types (for verification
+part) and then eliminate them (for using the original compiler), then
+our goal is achieved.
 
@@
 data structure, which is typically referred to as a tagless
-interpreter~\cite{}.  The benefit of deep embedding is that we get
+interpreter~\citep{PasalicTS02}.  The benefit of deep embedding is that we get
 full access to the syntactic structure of the embedded program.  This
 approach works beautifully for embedded languages with simple type
-systems. However, for embedded languages with dependent types,
-tagless interpreters are technically possible~\cite{}, but
-are extremely challenging to define and use in practice.
+systems. However, for embedded languages with dependent types, tagless
+interpreters are technically
+possible~\citep{CHAPMAN200921,10.1007/978-3-540-74464-1_7}, but are
+extremely challenging to define and use in practice.
 

### 2021-07-14 paper.tex
Full pass over paper, mostly cosmetic changes
Address some of my pet peeves

@@
 dependently typed language, properties can be encoded as types, and if
-the program typechecks, the property is guaranteed to hold.  If we can
-extend our existing language with dependent types (for verification
+the program typechecks, the property is guaranteed to hold.  If we
+extend our existing language with dependent types (for the verification
 part) and then eliminate them (for using the original compiler), then
@@
 To keep presentation as simple as possible, we have
-chosen a very simple target language to embed --- we use a subset
+chosen a simple target language to embed --- we use a subset
 of PostScript~\citep{postscript1999language}, a stack-based
@@
 
-This entire paper is a literate Agda script: all Agda code is typechecked
-and all the examples are run while generating the paper.
+This entire paper is a literate Agda script: all code is typechecked
+while generating the paper.
 
@@
 \citet{10.1145/3371071} introduce the Turnstile+ framework for
-building dependently typed embedded DSLs and very much
+building dependently typed embedded DSLs and
 shares the ideas advocated in this paper, suggesting that
@@
 is quite similar to our own. While they are able to formally
-reason about preservation of semantics (which we can't do yet), it is unclear
+reason about preservation of semantics (which we cannot do yet), it is unclear
 whether their approach scales to dependently-typed embedded languages.
@@
 language Agda.  We demonstrate the use of dependent types
-in order to enforce static properties of the embedded programs.
+to enforce static properties of the embedded programs.
 Concretely, we have demonstrated the approach by implementing

### 2021-07-15 paper.tex
Shrink to fit 12 pages :)
Use proper color for macro calls
Add section references to contributions

@@
 \usepackage[final]{listings}
+\lstset{
+    basicstyle=\small,
+}
 
@@
 \newunicodechar{ᵇ}{\ensuremath{{}^{b}}}
+\newunicodechar{✔}{\checkmark}
 
@@
 \newcommand{\AS}[1]{\AgdaString{#1}}
+\newcommand{\AMA}[1]{\AgdaMacro{#1}}
 
@@
 }
+
+\renewcommand{\AgdaCommentFontStyle}[1]{\textsf{#1}}
 
@@
 \item We demonstrate this approach by a concrete example: an embedding
-  of (a subset of the) PostScript language into Agda.
-
-\item We give several examples of PostScript programs written in our
-  embedding, and demonstrate how to statically enforce properties such
-  as the absence of stack underflows, termination, and functional
-  correctness.
+  of (a subset of the) PostScript language into Agda
+  (\secref{embedding}). We give several examples of PostScript
+  programs written in our embedding, and demonstrate how to statically
+  enforce properties such as the absence of stack underflows,
+  termination, and functional correctness.
 
 \item We show how to use reflection in Agda to implement an extractor
-  that translates code written in our embedding to plain PostScript.
+  that translates code written in our embedding to plain PostScript
+  (\secref{extraction}).
 
@@
   as automatically apply optimizations through the use of rewrite
-  rules.
+  rules (\secref{partial-evaluation}).
 \end{itemize}
 
+\begin{comment}
 The rest of this paper is structured as follows. \secref{background}
@@
 concludes.
+\end{comment}
 

### 2021-07-15 paper.tex
Do a (partial) pass over the paper

@@
 \newcommand{\AMA}[1]{\AgdaMacro{#1}}
+
+\newcommand{\hrefu}[2]{\href{#1}{\nolinkurl{#2}}}
 
@@
 It is often desirable to guarantee that a
-certain class of errors does not occur in the given program.  These
-error classes may include: dereferencing the null pointer, division
+certain class of errors does not occur in a given program.  These
+error classes may include: dereferencing a null pointer, division
 by zero, or out of bound indexing.  In most
-programming languages, the guarantees that can be checked
+programming languages, the properties that can be checked
 statically are limited.  Nevertheless, moving to a different
@@
 
-To keep presentation as simple as possible, we have
+To keep presentation as easy as possible to follow, we have
 chosen a simple target language to embed --- we use a subset
@@
 
-\paragraph{Metaprogramming} There is a large body of
+%\paragraph{Metaprogramming}
+There is a large body of
 work on metaprogramming facilities in various programming
 languages.  \citet{refl-masses} track the origins
-of metaprogramming to Smith's work on reflection in Lisp~\cite{refl-lisp}.
+of metaprogramming to Smith's work on reflection in Lisp \cite{refl-lisp}.
 Some prominent metaprogramming systems include
-MetaOcaml~\cite{metaocaml}, MetaML~\cite{metaml},
-reFlect~\cite{DBLP:journals/jfp/GrundyMO06},
-Template Haskell~\cite{sheard2002template},
-Racket~\cite{plt-tr1}, and various other Lisp/Scheme dialects.
+MetaOcaml \cite{metaocaml}, MetaML \cite{metaml},
+reFlect \cite{DBLP:journals/jfp/GrundyMO06},
+Template Haskell \cite{sheard2002template},
+Racket \cite{plt-tr1}, and various other Lisp/Scheme dialects.
 %
@@
 
-\paragraph{Embedding}
+%\paragraph{Embedding}
 Defining deep embeddings with static guarantees are a common application
@@
 we instead aim for a more lightweight approach.
-
+%
 \citet{deepshallow} propose to solve this problem with
@@
 a small deep embedding and leverage type classes in Haskell to define
-the rest of the language on top of that.  It would be interesting to see
-whether such an approach scales to dependently-typed embedded
-languages.
+the rest of the language on top of that.  However, so far this idea has not yet
+been applied to dependently typed embedded languages.
 
-\paragraph{Extraction}
+%\paragraph{Extraction}
 The Coq proof assistant is equipped with extraction
@@
 
-\paragraph{Dependently typed metaprogramming}
+%\paragraph{Dependently typed metaprogramming}
 Several dependently-typed languages are equipped with metaprogramming

### 2021-07-15 paper.tex
use wrapfig to save more space
Kill hanging lines, center fig 1

@@
 \usepackage{epsf}
+\usepackage{wrapfig}
 
@@
   Most existing programming languages provide little support to formally
-  state and prove properties about programs.  Adding such capabilities is far
+  state and prove properties about programs.  Ad\-ding such capabilities is far
   from trivial, as it requires significant re-engineering of the existing
   compilers and tools.  This paper proposes a novel technique to write
-  correct-by-construction programs in languages without built-in verification
+  correct-by-con\-struc\-tion programs in languages without built-in verification
   capabilities, while maintaining the ability to use existing tools.
@@
   express all three steps in a single language that supports
-  both dependent types and reflection. Essentially,
+  both dependent typ\-es and reflection. Essentially,
   this allows us to express a program, its formal properties, and
@@
 Some prominent metaprogramming systems include
-MetaOcaml \cite{metaocaml}, MetaML \cite{metaml},
+MetaOcaml \cite{metaocaml}, Me\-taML \cite{metaml},
 reFlect \cite{DBLP:journals/jfp/GrundyMO06},

### 2021-07-16 paper.tex
bugfix in sierpinski, add colors to urls and citations
Minor fixes in the intro (and bib file)

@@
-\documentclass[sigplan,anonymous,review]{acmart}
+\documentclass[sigplan,screen,anonymous,review]{acmart}
 
@@
 }
+\usepackage{microtype}
 
@@
 of the host language.  Programs in a shallowly-embedded language are programs
-in the host language, so they are immediately runnable; types of the embedded
+in the host language, so they are immediately run\-nable; types of the embedded
 language are types of the host language, so the host typechecker is used for
@@
 \item We demonstrate this approach by a concrete example: an embedding
-  of (a subset of the) PostScript language into Agda
+  of a subset of the PostScript language into Agda
   (\secref{embedding}). We give several examples of PostScript

### 2021-09-21 paper.tex
Restore 12 pages limit
Trivial fixes.

@@
-\documentclass[sigplan,screen,anonymous,review]{acmart}
+\documentclass[sigplan,screen]{acmart}
 
@@
   that had to go without them so far.
-
-%  Dependently-typed host languages empower users to verify a wide
-%  range of properties of embedded languages and
-%  programs written in them.  Designers of such embedded languages
-%  are faced with a difficult choice between using a shallow or a deep
-%  embedding.  The former is easier to use because the entire
-%  infrastructure of the host langauge is immediately
-%  available. Meanwhile, the latter gives full access to the
-%  structure of embedded programs, but is difficult to use in practice,
-%  especially when the embedded language is itself dependently typed.
-%
-%  The main insight presented in this paper is that the choice between
-%  shallow and deep embedding can be eliminated by working in a host
-%  language with reflection capabilities: we start from a shallow
-%  embedding that can use all libraries and tools of the host language,
-%  and later use reflection to expose the deep structure of the
-%  embedded programs.
-%  %
-%  Concretely, we apply this technique to embed three
-%  programming languages --- Kaleidoscope, SaC, and (a subset of) APL ---
-%  into the dependently typed theorem prover Agda, using dependent
-%  types to statically enforce several properties of interest.  We then
-%  use Agda's reflection capabilities to extract the embedded programs
-%  back into the original language, so that the existing toolchain can
-%  be leveraged.
-%  %
-%  In this process, statically verified properties of the host language
-%  are mapped onto runtime checks in the target language, allowing
-%  extracted programs to interact safely with existing code.
-%  %
-%  Finally, we demonstrate the feasibility of our approach with the
-%  implementation and extraction of a convolutional neural network in
-%  our embedding of APL.\@
-
-%  Most practical programming languages of today prioriterise
-%  performance and usability over safety.  As a result, many
-%  useful correctness invariants cannot be expressed in the given
-%  language, typically due to weakness of the chosen type system.
-%  We are looking for non-invasive verification: how to
-%  guarantee a property about the given program without any modifications
-%  to the language toolchain.
-%
-%  We study an approach when a programming language is shallowly
-%  embedded into a theorem prover, and dependent types are used
-%  to encode the properties of interest.   Bringing the embedded
-%  langauge and the properties back into the original language
-%  is the main focus of this work.  We propose a generic extraction
-%  mechanism that we implement in Agda that is based on the concept
-%  of reflection.  Making it possible to translate any shallowly
-%  embedded language in the target backend without the necessity
-%  to modify the theorem prover.
-%  We demonstrate our framework in action by encoding three languages:
-%  Kaleidoskope, SaC and a subset of APL, culminating in the implementation
-%  and extraction of a convolutional neural network.
 \end{abstract}
@@
 Some prominent metaprogramming systems include
-MetaOcaml \cite{metaocaml}, Me\-taML \cite{metaml},
+MetaOCaml \cite{metaocaml}, Me\-taML \cite{metaml},
 reFlect \cite{DBLP:journals/jfp/GrundyMO06},
@@
 which extracts functional code from Coq proofs (or programs).  The
-default target language is Ocaml, but a few other options were added
+default target language is OCaml, but a few other options were added
 recently.
@@
 
+\pagebreak
 %% Bibliography

### 2021-09-22 paper.tex
Acknowledge grants.
Adding copyright.
        This will need further updates after I'll figure out the deal
        with the uninversity paying for open access.

@@
 \documentclass[sigplan,screen]{acmart}
+
+%%% FIXME I am didnt update publishing rights yet, as I am waiting for
+%%% the university to comeback with the payment instructions for open access.
+
+%%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
+%%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
+%%% Then please update our instructions page and copy-and-paste the NEW commands into your article.
+%%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
+%%%
+%%% The following is specific to GPCE '21 and the paper
+%%% 'Extracting the Power of Dependent Types'
+%%% by Artjoms Šinkarovs and Jesper Cockx.
+%%%
+\setcopyright{rightsretained} % XXX ENSURE THAT THIS VALUE IS CORRECT!!!
+\acmPrice{15.00}
+\acmDOI{10.1145/3486609.3487201}
+\acmYear{2021}
+\copyrightyear{2021}
+\acmSubmissionID{splashws21gpcemain-p18-p}
+\acmISBN{978-1-4503-9112-2/21/10}
+\acmConference[GPCE '21]{Proceedings of the 20th ACM SIGPLAN International Conference on Generative Programming: Concepts and Experiences}{October 17--18, 2021}{Chicago, IL, USA}
+\acmBooktitle{Proceedings of the 20th ACM SIGPLAN International Conference on Generative Programming: Concepts and Experiences (GPCE '21), October 17--18, 2021, Chicago, IL, USA}
 
@@
 
-\pagebreak
+\begin{acks}
+We would like to thank reviewers for their constructive suggestions.
+This work is supported by the \grantsponsor{EPSRC}{Engineering and Physical
+Sciences Research Council}{https://epsrc.ukri.org} through the grant
+\grantnum{EPSRC001}{EP/N028201/1}.
+\end{acks}
+
+%\pagebreak
 %% Bibliography

### 2021-09-23 paper.tex
Addressing reviewer's comments (hopefully).

@@
   in that embedding, and we use dependent types to guarantee
-  correctness properties of interest.  Thirdly, we
+  correctness properties of interest within the embedding.  Thirdly, we
   extract a program written in the original language,
@@
 There are two common approaches to language embedding: deep and shallow.
-A shallow embedding defines a language in terms of functions and operators
-of the host language.  Programs in a shallowly-embedded language are programs
+Shallow embeddings define languages in terms of functions and operators
+of the host language.  Programs in shallowly-embedded languages are programs
 in the host language, so they are immediately run\-nable; types of the embedded
@@
 data structure, which is typically referred to as a tagless
-interpreter~\citep{PasalicTS02}.  The benefit of deep embedding is that we get
+interpreter~\citep{PasalicTS02,Tagless}.  The benefit of deep embedding is that we get
 full access to the syntactic structure of the embedded program.  This
@@
 possible~\citep{CHAPMAN200921,10.1007/978-3-540-74464-1_7}, but are
-extremely challenging to define and use in practice.
+\emph{extremely challenging} to define and use in practice.
+For further details on why exactly this is challenging refer to the
+first section of~\cite{10.1145/1863495.1863497} where the problem is meticulously
+explained by~\citeauthor{10.1145/1863495.1863497}.
 
@@
 %
-While the embedding may look simplistic, this simplicity
-allows us to focus on the essence of the proposed approach and avoid unrelated
-technical details.  We ensure that many verification patterns found in bigger
-languages are covered, showing that the proposed approach scales.
+This paper is not concerned with practical application of
+the presented embedding, so we consciously keep the setup as minimal as
+possible.  It is powerful enough to motivate the use of dependent
+types \emph{within} the embedding, and to demonstrate the essence
+of the proposed approach.  While we include many patterns found
+in bigger languages, we leave the scaling of the approach out
+of the scope of the paper referring the interested reader
+to~\cite{DBLP:journals/corr/abs-2105-10819} where we work
+out such generalisations.
 
-This entire paper is a literate Agda script: all code is typechecked
-while generating the paper.
+% While the embedding may look simplistic, this simplicity
+% allows us to focus on the essence of the proposed approach and avoid unrelated
+% technical details.  We ensure that many verification patterns found in bigger
+% languages are covered, showing that the proposed approach scales.
+
+This entire paper is a literate Agda script\footnote{The sources of the paper
+can be found at \url{https://github.com/ashinkarov/agda-stacklang}.}:
+all code is typechecked while generating the paper.
 
@@
 \end{itemize}
+
+When dealing with embedded languages that are extracted and run
+by external compilers, we are fundamentally bound to have potential
+semantic mismatches.  Semantics of the embedding may differ from the
+semantics implemented by external compilers, as well as translation from the
+embedding back to external compilers can be faulty.  Real world compilers
+face exactly the same problem: how do you know that generated assembly
+code is doing what you think it is doing?  Addressing such problems
+requires fully verified toolchains that run on a verified hardware.
+Enormous research efforts are spent towards addressing this
+problem, but this is beyond the scope of this paper.
 
@@
 Coq \cite{metacoq}, and Agda \cite{agda-refl}.  All of these
-implement a similar API as described in this paper.  This is
-reassuring, as it means our proposed approach is immediately portable
-into many other contexts.
+implement a similar API as described in this paper.  This
+means that our proposed approach is immediately implementable
+in these languages --- one simply needs to adjust extraction
+to the particular API.
 \citet{10.1145/3371071} introduce the Turnstile+ framework for

### 2021-09-24 paper.tex
Updating ACM copyright after filling-in right transfer.
A few additions/tweaks to the parts addressing reviewer comments

@@
 
-%%% FIXME I am didnt update publishing rights yet, as I am waiting for
-%%% the university to comeback with the payment instructions for open access.
-
-%%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
-%%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
-%%% Then please update our instructions page and copy-and-paste the NEW commands into your article.
-%%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
-%%%
-%%% The following is specific to GPCE '21 and the paper
-%%% 'Extracting the Power of Dependent Types'
-%%% by Artjoms Šinkarovs and Jesper Cockx.
-%%%
-\setcopyright{rightsretained} % XXX ENSURE THAT THIS VALUE IS CORRECT!!!
-\acmPrice{15.00}
-\acmDOI{10.1145/3486609.3487201}
+% Updating copyright info, as per latest ACM email
+% after copyright transfer.
+\copyrightyear{2021}
 \acmYear{2021}
-\copyrightyear{2021}
-\acmSubmissionID{splashws21gpcemain-p18-p}
-\acmISBN{978-1-4503-9112-2/21/10}
+\setcopyright{rightsretained}
 \acmConference[GPCE '21]{Proceedings of the 20th ACM SIGPLAN International Conference on Generative Programming: Concepts and Experiences}{October 17--18, 2021}{Chicago, IL, USA}
 \acmBooktitle{Proceedings of the 20th ACM SIGPLAN International Conference on Generative Programming: Concepts and Experiences (GPCE '21), October 17--18, 2021, Chicago, IL, USA}
+\acmDOI{10.1145/3486609.3487201}
+\acmISBN{978-1-4503-9112-2/21/10}
 
@@
   extract a program written in the original language,
-  so that it can be used by the existing compilers and tools.
+  so it can be used with existing compilers and tools.
 
@@
 possible~\citep{CHAPMAN200921,10.1007/978-3-540-74464-1_7}, but are
-\emph{extremely challenging} to define and use in practice.
-For further details on why exactly this is challenging refer to the
-first section of~\cite{10.1145/1863495.1863497} where the problem is meticulously
-explained by~\citeauthor{10.1145/1863495.1863497}.
+\emph{extremely} challenging to define and use in practice.
+%
+In particular, the mutual dependencies between the syntax for
+contexts, types, and terms and the definitional equality of the object
+language require large and complex inductive-inductive or
+inductive-recursive to encode
+\cite{10.1007/978-3-540-74464-1_7,CHAPMAN200921,10.1145/1863495.1863497}.
+%
+Even if we manage to give a faithful deep embedding of the language,
+actually writing programs in it is still an arduous task compared
+to writing programs in the meta-language directly.
 
@@
 semantic mismatches.  Semantics of the embedding may differ from the
-semantics implemented by external compilers, as well as translation from the
-embedding back to external compilers can be faulty.  Real world compilers
+semantics implemented by external compilers, and the translation from the
+embedding back to external compilers can also be faulty.  Real world compilers
 face exactly the same problem: how do you know that generated assembly
@@
 Enormous research efforts are spent towards addressing this
-problem, but this is beyond the scope of this paper.
+problem~\cite{Leroy09,KumarMNO14}, but this is beyond the scope of this paper.
 

### 2021-09-24 paper.tex
Add my orcid + acknowledge NWO

@@
                                          %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\orcid{0000-0003-3862-4073}              %% \orcid is optional
 \affiliation{%
@@
 We would like to thank reviewers for their constructive suggestions.
-This work is supported by the \grantsponsor{EPSRC}{Engineering and Physical
-Sciences Research Council}{https://epsrc.ukri.org} through the grant
-\grantnum{EPSRC001}{EP/N028201/1}.
+This work is supported by the \grantsponsor{EPSRC}{Engineering and
+Physical Sciences Research Council}{https://epsrc.ukri.org} through
+the grant \grantnum{EPSRC001}{EP/N028201/1}, and the
+\grantsponsor{NWO}{Dutch Research Council}{https://www.nwo.nl/en}
+through the grant VI.Veni.202.216.
 \end{acks}

### 2021-09-24 paper.tex
Minor typo

@@
 language require large and complex inductive-inductive or
-inductive-recursive to encode
+inductive-recursive types to encode
 \cite{10.1007/978-3-540-74464-1_7,CHAPMAN200921,10.1145/1863495.1863497}.

### 2021-09-24 paper.tex
First attempt on trimming the paper back into 12 pages.
        * removed figure from the intro (for now)
        * removed most of the secion about monad (needs to be
          reintroduced gently)
        * aggressively wrapped figures that are < 2/3 column
        * redefined \paragraph to avoid vertical spacing

@@
 \newcommand{\eqnref}[1]{\hyperref[#1]{(\ref{#1})}}
+
+\renewcommand{\paragraph}[1]{\textbf{\textit{#1. }}}
 
@@
 
-\input{latex/intro-example}
+% \input{latex/intro-example}
 
@@
 
-This entire paper is a literate Agda script\footnote{The sources of the paper
+This entire paper is a \emph{literate Agda} script\footnote{The sources of the paper
 can be found at \url{https://github.com/ashinkarov/agda-stacklang}.}:
@@
 
-\paragraph{Contributions.}
+Our \textit{\textbf{main contributions}} are as follows:
+
 \begin{itemize}
@@
 
-When dealing with embedded languages that are extracted and run
+When dealing with embedded languages that are ex\-trac\-ted and run
 by external compilers, we are fundamentally bound to have potential

### 2021-09-27 paper.tex
Further trimming.
        * Added (back) compressed  paragraph about monads.
        * Got rid of a^2+b^2 example from the intro.
        * Important: Globally reduced spacing for wrapfigures
        * Important: Macro that turns paragraph environment to
          use less spacing (off now).
        * Important: Reduced spacing around agda code blocks.
          Unfortunately in agda.sty for now, maybe one can do this
          in papart.tex.
99% compressed (two lines remain!)
        * Reduced spacing around mathpar
        * redefined \paragraph to \noindentparagraph.

@@
 
-\renewcommand{\paragraph}[1]{\textbf{\textit{#1. }}}
+%\renewcommand{\paragraph}[1]{\textbf{\textit{#1. }}}
+\setlength{\intextsep}{0pt}
+%\AgdaNoSpaceAroundCode{}
+\renewcommand\paragraph{\noindentparagraph}
+
+% This trick is taken from here:
+% https://tex.stackexchange.com/questions/346444/mathpartir-abovedisplayskip-and-belowdisplayskip-affect-placement-of-side-con
+
+\usepackage{etoolbox}
+
+\newcommand{\zerodisplayskips}{%
+  \setlength{\abovedisplayskip}{2pt}%
+  \setlength{\belowdisplayskip}{2pt}%
+  \setlength{\abovedisplayshortskip}{2pt}%
+  \setlength{\belowdisplayshortskip}{2pt}}
+\appto{\normalsize}{\zerodisplayskips}
+\appto{\small}{\zerodisplayskips}
+\appto{\footnotesize}{\zerodisplayskips}
 
@@
 There are two common approaches to language embedding: deep and shallow.
-Shallow embeddings define languages in terms of functions and operators
+Shallow embeddings define lan\-gua\-ges in terms of functions and operators
 of the host language.  Programs in shallowly-embedded languages are programs
@@
 %
-\figref{sqsum} shows an example of a program
-written in our embedding and the PostScript code produced by
-our extractor.
+%\figref{sqsum} shows an example of a program
+%written in our embedding and the PostScript code produced by
+%our extractor.
 %
@@
 % languages are covered, showing that the proposed approach scales.
-
-This entire paper is a \emph{literate Agda} script\footnote{The sources of the paper
-can be found at \url{https://github.com/ashinkarov/agda-stacklang}.}:
-all code is typechecked while generating the paper.
 
@@
   evaluation by making use of normalisation in the host language. This
-  allows us to use functions of the host language as macros, as well
-  as automatically apply optimizations through the use of rewrite
-  rules (\secref{partial-evaluation}).
+  allows us to use functions of the host language as macros,
+  and optimise prorgams via rewrite rules (\secref{partial-evaluation}).
+
+  %as well
+  %as automatically apply optimizations through the use of rewrite
+  %rules (\secref{partial-evaluation}).
 \end{itemize}
@@
 problem~\cite{Leroy09,KumarMNO14}, but this is beyond the scope of this paper.
+
+This entire paper is a \emph{literate Agda} script\footnote{The sources of the paper
+can be found at \url{https://github.com/ashinkarov/agda-stacklang}.}:
+all code is typechecked while generating the paper.
 

### 2021-09-27 paper.tex
Fix overfull lines
Pass for small optimizations for space and layout

@@
 \usepackage{microtype}
+
+% Fixing overfull lines
+\emergencystretch=10.0pt % default 0.0pt
+\hyphenpenalty=500 % default 50
+\exhyphenpenalty=500 % default 50
 
@@
 Shallow embeddings define lan\-gua\-ges in terms of functions and operators
-of the host language.  Programs in shallowly-embedded languages are programs
+of the host language.  Programs in shallowly embedded languages are programs
 in the host language, so they are immediately run\-nable; types of the embedded
 language are types of the host language, so the host typechecker is used for
-checking embedded programs.  The downside of this approach is that we cannot
-easily get the representation of embedded programs.
+checking embedded programs. However we cannot easily get the syntactic
+representation of a shallowly embedded program.
 
@@
 %
-While the fact that this is possible is impressive in theory, the resulting
-encodings are often difficult to use in practice. In this paper
+While the this is impressive in theory, the resulting
+encodings are difficult to use in practice. In this paper
 we instead aim for a more lightweight approach.
@@
 \citet{deepshallow} propose to solve this problem with
-a combination of deep and shallow embeddings.  Their idea is to define
-a small deep embedding and leverage type classes in Haskell to define
+a combination of deep and shallow embeddings, by using
+a small deep embedding and leveraging type classes in Haskell to define
 the rest of the language on top of that.  However, so far this idea has not yet
@@
 %
-While it would be possible to implement extractors presented in this
-paper as Coq plugins or Agda backends, conceptually they are more
-heavyweight.  Our extractors and programs can (in principle)
+As a consequence, our extractors and programs can (in principle)
 communicate with each other. In addition, as they are just Agda programs, they can
@@
 Coq \cite{metacoq}, and Agda \cite{agda-refl}.  All of these
-implement a similar API as described in this paper.  This
-means that our proposed approach is immediately implementable
-in these languages --- one simply needs to adjust extraction
-to the particular API.
+implement a similar API as described in this paper, and hence could be
+used to implement our approach --- one simply needs to adjust
+extraction to the particular API.
 \citet{10.1145/3371071} introduce the Turnstile+ framework for
@@
 This allows us to enjoy simplicity of writing programs
-in shallow embedding.  Yet we keep full access to the internal
-structure of the embedded programs, so that we could translate
+in shallow embedding, while keeping the ability to translate
 them into the original language.
@@
 APL (Array Programming Language), another array language with
-heavily overloaded syntax.
-For more details on these examples,
-refer to~\citet{DBLP:journals/corr/abs-2105-10819}.
+heavily overloaded syntax~\cite{DBLP:journals/corr/abs-2105-10819}.
 
@@
 
-\paragraph{Future work}
-While our approach is flexible and powerful, it is
-certainly not a magic bullet. A lot of things
-need to be implemented for each target language: the embedding,
-the encoding of the properties under consideration as types,
-and the extractor.  For enforcing more complex properties, a number
-of proofs needs to be written explicitly.
+\paragraph{Future work} While our approach is flexible, it is not a
+magic bullet. Right now, we cannot yet guarantee that the extracted
+code preservers the semantics of the original implementation. While we
+rarely see fully-verified compiler backends in the real world, our
+approach is very close to enabling this.  In future work, we would
+like to give a formal semantics of the reflected language and the
+proof that reflected programs respect it.  While this is non-trivial,
+a system like Agda could do it in principle.
 
-Right now, we cannot yet guarantee that the
-extracted code preservers the semantics of the original
-implementation. While we rarely see
-fully-verified compiler backends in the real world,
-our approach is very close to enabling this.
-We would need a formal semantics of the reflected language
-and the proof that reflected programs respect it.
-While this is non-trivial, a system like Agda could do
-this in principle.
+%\paragraph{Future work}
+%While our approach is flexible and powerful, it is
+%certainly not a magic bullet. A lot of things
+%need to be implemented for each target language: the embedding,
+%the encoding of the properties under consideration as types,
+%and the extractor.  For enforcing more complex properties, a number
+%of proofs needs to be written explicitly.
 

### 2021-09-27 paper.tex
Don't use noindentparagraph
I find it rather ugly
Add keywords & CSS concepts

@@
 %\AgdaNoSpaceAroundCode{}
-\renewcommand\paragraph{\noindentparagraph}
+%\renewcommand\paragraph{\noindentparagraph}
 
@@
 \end{abstract}
+
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011006.10011041.10011047</concept_id>
+<concept_desc>Software and its engineering~Source code generation</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10011006.10011041.10011046</concept_id>
+<concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10010940.10010992.10010998.10010999</concept_id>
+<concept_desc>Software and its engineering~Software verification</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10011006.10011050.10011017</concept_id>
+<concept_desc>Software and its engineering~Domain specific languages</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[300]{Software and its engineering~Source code generation}
+\ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}
+\ccsdesc[300]{Software and its engineering~Software verification}
+\ccsdesc[300]{Software and its engineering~Domain specific languages}
+
+\keywords{embedded languages, program extraction, program verification, dependent types, reflection, Agda}
+
 \maketitle

### 2021-09-27 paper.tex
Fixing warnings produced by submission
Make veritucal spacing a bit nicer.

@@
 \setcopyright{rightsretained}
+\acmPrice{}
 \acmConference[GPCE '21]{Proceedings of the 20th ACM SIGPLAN International Conference on Generative Programming: Concepts and Experiences}{October 17--18, 2021}{Chicago, IL, USA}
@@
 
-\usepackage{todonotes}
+%\usepackage{todonotes}
 
@@
 % Fixing overfull lines
-\emergencystretch=10.0pt % default 0.0pt
+\emergencystretch=2pt % default 0.0pt
 \hyphenpenalty=500 % default 50
@@
 %% Title information
-\title[]{Extracting The Power of Dependent Types}
+\title[]{Extracting the Power of Dependent Types}
 
@@
 Physical Sciences Research Council}{https://epsrc.ukri.org} through
-the grant \grantnum{EPSRC001}{EP/N028201/1}, and the
+the grant
+
+\noindent
+\grantnum{EPSRC001}{EP/N028201/1}, and the
 \grantsponsor{NWO}{Dutch Research Council}{https://www.nwo.nl/en}
-through the grant VI.Veni.202.216.
+through the grant~VI.Veni.202.216.
 \end{acks}
